---
title: "Compile Poudre Water Chemistry"
author: "Sam Struthers- CSU ROSSyndicate"
date: "`r Sys.Date()`"
output: html_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("scripts/00_setup.R")

```

# Import Water Chemistry Dataset
 If new data is published, the DOI and data version need to match the most recent publication on Zenodo

```{r pressure, echo=FALSE}

#discrete sample data and location metadata from most recent pub

DOI = "11110885"
data_version = "v2024.04.16b"


source("scripts/grab_zenodo_data.R")

# all_params <- chem_units%>%
#   pull(simple)
```

# Grab Sensor Data
Sensor data comes in two sources: Livestreamed data to HydroVu and data which is sent via radio network (operated by WET and Novastar)

## Grab HydroVu Data

### Setup

```{r}

# Only grab sites in the chemistry dataset that have a sonde
sites_oi <- most_recent_chem%>%
  filter(!is.na(WQ_sensor)& site_code %nin% c("PNF", "PBR", "PFAL", "ELC", "SFM", "legacy") )%>%
  pull(site_code)%>%
  unique()

sites <- c("JOEI", "CBRI", "CHD", 
           # Not yet uploading to HydroVu: "PFAL", "SFM", "LBEA", "PENN", 
           "PBD", "Lincoln", "Timberline", "Prospect","Boxelder",  "Archery", 
           "BoxCreek", "SpringCreek")

#Just for testing
#sites_oi <- c("prospect", "archery")
# Read in HydroVu Creds from credentials.yml
hv_creds <- read_yaml("scripts/api_pull/credentials.yml")
hv_token <- hv_auth(client_id = as.character(hv_creds["client"]),
                      client_secret = as.character(hv_creds["secret"]),
                      url = "https://www.hydrovu.com/public-api/oauth/token")

```

### Set start and end time for API grab
```{r}
# Grab sample + sensors started in late August 2023, we will grab from the start of the month to be safe
start_dt <- ymd_hms("2023-08-01 00:00:00", tz = "MST")
# End date is the day after the last sample was grabbed
end_dt <- Sys.Date()

```
## Pull HydroVu Api

```{r}
source("scripts/api_pull/api_puller.R")
source("scripts/api_pull/hv_locations_all.R")
source("scripts/api_pull/hv_getdata_id.R")
source("scripts/api_pull/munge_api_data.R")
#grab data
# grab <- map(sites_oi,~ api_puller(site = .x,
#                  start_dt = start_dt,
#                  end_dt = end_dt, 
#                  api_token = hv_token,
#                  dump_dir = "data/api"))


#import data from API folder
sensor_data <- munge_api_data(api_path = "data/api")

```
## Pull in SFM data from GH

```{r}
# download directly from a GH repository named SamStruthers/ross_sfm, this is a public repo where I backup the livestreaming data from SFM

# Open a connection to the URL
con <- url("https://github.com/SamStruthers/Sonde_livestream_backup/raw/main/data/archive/SF_data_archive2023.RDS", "rb")
# Read the RDS file directly from the connection
sfm_2023_data <- readRDS(gzcon(con))%>%
  distinct()

# Open a connection to the URL
con <- url("https://github.com/SamStruthers/Sonde_livestream_backup/raw/main/data/SF_data_archive2024.RDS", "rb")
# Read the RDS file directly from the connection
sfm_2024_data <- readRDS(gzcon(con))%>%
  distinct()

sfm_data <- bind_rows(sfm_2023_data, sfm_2024_data)%>%
  mutate(site = "sfm", 
#fixing parameter names
         parameter = case_when( Measurement == "Temperature" ~ "ross_temp",
                                Measurement == "pH" ~ "ross_pH",
                                Measurement == "DO" ~ "ross_DO",
                                Measurement == "DO_Sat" ~ "ross_DO_sat",
                                Measurement == "Actual_Conductivity" ~ "ross_turb",
                                Measurement == "Specific_Conductivity" ~ "ross_spec_cond",
                                Measurement == "Turbidity" ~ "ross_turb",
                                Measurement == "Chl-a" ~ "ross_chla",
                                Measurement == "FDOM" ~ "ross_FDOM",
                                Measurement == "Depth" ~ "ross_depth",
                                TRUE ~ NA))%>%
        
  select(site, DT, DT_round, value = Value, parameter)%>%
  distinct()
  

  write_csv(sfm_data, "data/sfm_data.csv")
  rm(sfm_2023_data, sfm_2024_data, con)
```



## Tidying hydrovu data

```{r}

long_sensor_data <- sensor_data%>%
  filter(!is.na(value))%>%
  #rename parameter for sites where sondes are colocated
  left_join(sensor_meta%>% select(param_sonde, param_common), by = c("parameter" = "param_sonde"))%>%
  mutate(parameter = case_when(grepl("virridy", name, ignore.case = TRUE) ~ paste0("virridy_", param_common),
                           TRUE ~ paste0("ross_", param_common)))%>% 
  select(site, DT, DT_round,  value, parameter)%>%
  rbind(sfm_data)

wide_sensor_data <- long_sensor_data%>%
  pivot_wider(names_from = parameter, values_from = value, id_cols = c(site, DT))


```

## Joining datasets

```{r}

sites <- c("JOEI", "CBRI", "CHD", 
           "PFAL", "SFM", "LBEA", "PENN", 
           "PBD", "lincoln", "timberline", "prospect","boxelder",  "archery", 
           "boxcreek", "springcreek")


grabs_at_sensors <- most_recent_chem%>%
  #remove data before sensor data starts
  filter( site_code %in% sites, DT_mst >= min(wide_sensor_data$DT))%>%
  #lower case to match sensor data
  mutate(site_code = tolower(site_code), 
         grab_dt = round_date(DT_mst, "15 minutes"))%>%
  select(site_code, grab_dt, Turbidity:SO4, Field_DO_mgL,Field_Cond_µS_cm,Field_Temp_C )


  merge_grab_sensor <- function( grab_dt, site_select){
    # convert to date to filter sensor data to only data on same date as sample
    grab_date = as.Date(grab_dt, tz = "MST")
    
    # grab data from 4 data points after a site visit
      after_data <- wide_sensor_data %>%
        #create date for match up
        mutate(date = as.Date(DT, tz = "MST")) %>%
        filter(grepl(site_select, site, ignore.case = TRUE) & date == grab_date & DT > grab_dt )%>%
        dplyr::arrange(DT)%>%
        # grab the data 4 data points after a site visit
        dplyr::slice_head(n = 4)
      
        #grab data before grab sample time, this is to be used for when a sonde was pulled, essentially reference data prior to the sonde being pulled rather than after
        before_data <- wide_sensor_data %>%
          #create date for match up
          mutate(date = as.Date(DT, tz = "MST")) %>%
          filter(grepl(site_select, site, ignore.case = TRUE) & date == grab_date & DT < grab_dt )%>%
          dplyr::arrange(DT)%>%
          # grab the data missing data points from after_data to make sure we have 4 data points
          dplyr::slice_tail(n = (4- nrow(after_data)) )
    
        
        #bind data together
        select_sensor_data <- bind_rows(before_data, after_data)
        #create a nested data frame for the sensor data
        nest_data <- select_sensor_data%>%nest()
        
        if(nrow(select_sensor_data) == 0){
          
          select_sensor_data%>%
       # take the median of all sensor values
       summarise_if(is.numeric, median, na.rm = TRUE)%>%
       #add site and grab_dt back in
       mutate(site = site_select, grab_dt = grab_dt)
        }else{
       select_sensor_data%>%
       # take the median of all sensor values
       summarise_if(is.numeric, median, na.rm = TRUE)%>%
       #add site and grab_dt back in
       mutate(site = site_select, grab_dt = grab_dt)%>%
            #add the nested data for future reference
            bind_cols(nest_data)
      
        }
  }
    
      
  correlated_df <- grabs_at_sensors%>%
    #grab only dt and site from discrete samples for calculate avg fucn
    select(grab_dt, site_select = site_code)%>%
    # map over calc avg fucn
    pmap_dfr(., merge_grab_sensor)%>%
    filter(rowSums(!is.na(select(., -grab_dt, -site))) > 0)%>%
    # join with lab data by site and grab dt
    left_join(grabs_at_sensors, by = c("grab_dt", "site" = "site_code"))
  
  tidy_correlated_df <- correlated_df%>%
    #remove sites with no sensor data
     filter(map_lgl(data, ~!is.null(.)))%>%
    select(site, grab_dt, data, everything())%>%
    arrange(site, grab_dt)
  
write_csv(correlated_df, "data/combined/correlated_data.csv")
```





