---
title: "Compile Poudre Water Chemistry"
author: "Sam Struthers- CSU ROSSyndicate"
date: "`r Sys.Date()`"
output: html_document
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
source("scripts/00_setup.R")

```

# Import Water Chemistry Dataset

If new data is published, the DOI and data version need to match the most recent publication on Zenodo

Current release: [link](https://zenodo.org/records/12752311)

```{r loading_data, echo=FALSE}

#discrete sample data and location metadata from most recent pub


DOI = "12752311"
data_version = "v.2024.07.16"

source("scripts/functions/grab_zenodo_data.R")

# all_params <- chem_units%>%
#   pull(simple)

rm(chem_units)
```

## Pull in flow data

This chunk will pull in the hourly median for flow data (cfs) from CDWR or USGS gauges for each samples where datetimes are available.

```{r}

gather_q_for_sample <- function(Date, site_code, DT_mst, flow_gauge_source, flow_gauge_id) {
  #recompile info into a tibble
  meta <- tibble(site_code = site_code, DT_mst = DT_mst, flow_gauge_source = flow_gauge_source, flow_gauge_id = flow_gauge_id, date = Date) 
  
  #skip sites not from USGS or CDWR
        if(meta$flow_gauge_source %nin% c("USGS", "CDWR")) {
          return(NULL)
        }
  #only run thru days where we have a datetime
  if(is.na(DT_mst)){
    return(NULL)
  }
  
        # JOEI gauge down starting 10/17/2023
  if(meta$site_code == "JOEI" & DT_mst >= ymd_hm("2023-10-17 00:00", tz = "MST")){
      return(NULL)
  }


  q_data <- tryCatch({
    if (meta$flow_gauge_source == "USGS") {
      readNWISuv(siteNumbers = meta$flow_gauge_id,
                 startDate = meta$date - days(1),
                 endDate = meta$date + days(1),
                 parameterCd = "00060", tz = "America/Denver") %>%
        distinct() %>%
        mutate(DT_mst = with_tz(dateTime, tzone = "MST"),
               source = "USGS") %>%
        select(site = site_no, DT_mst, q_cfs = X_00060_00000, flag = X_00060_00000_cd)
    } else if (meta$flow_gauge_source == "CDWR") {
      get_telemetry_ts(
        abbrev = meta$flow_gauge_id,
        parameter = "DISCHRG",
        start_date = meta$date - days(1),
        end_date = meta$date + days(1),
        timescale = "raw",
        include_third_party = TRUE) %>%
        distinct() %>%
        mutate(DT_mst = force_tz(datetime, tzone = "MST"),
               flag = case_when(!is.na(flag_a) & !is.na(flag_b) ~ paste0(flag_a, ",", flag_b),
                                is.na(flag_a) & !is.na(flag_b) ~ flag_b,
                                is.na(flag_b) & !is.na(flag_b) ~ flag_a,
                                TRUE ~ NA_character_),
               source = "DWR") %>%
        select(site = abbrev, DT_mst, q_cfs = meas_value, flag, source)
    }
  }, error = function(e) {
    message("Data unavailable for ", meta$site_code, " on ", meta$date)
    return(NULL)
    })
 
  if(is.null(q_data)) {
   return(NULL) 
  }

    sample_hour <- round_date(meta$DT_mst, "hour")
        
    hourly_q_cfs <- q_data %>%
      mutate(DT_mst = floor_date(DT_mst, "hour")) %>%
      filter(DT_mst == sample_hour) %>%
      summarise(hourly_q_cfs = median(q_cfs, na.rm = TRUE)) %>%
      pull(hourly_q_cfs)

  
  final_q <- tibble(Date, DT_mst,flow_gauge_id, hourly_q_cfs)
  
  return(final_q)

  }

q_for_chem <- most_recent_chem %>%
  select(Date, site_code, DT_mst, flow_gauge_source, flow_gauge_id)%>%
  pmap(gather_q_for_sample)%>%
  bind_rows()%>%
  filter(!is.na(hourly_q_cfs))


 most_recent_chem <- most_recent_chem%>%
  left_join(q_for_chem%>%select(hourly_q_cfs, DT_mst, flow_gauge_id), by = c("DT_mst", "flow_gauge_id"))


```



# Pull in CSU verified data

Sensor data was pulled in by ROSS via the workflow in the repository `poudre_sonde_network/virridy/non_targets_workflow`: [link](https://github.com/rossyndicate/poudre_sonde_network)

After being passed through automatic QA/QC, the data was then verified by ROSS team members to remove additional erronous data. The workflow for this step is available in the repository `poudre_sonde_network/virridy/verification_workflow`

Contact Katie Willi or Juan De La Torre for QAQC/verification workflow details. 

```{r}


# directory with verified data
ver_dir <- "data/verified_sensor_data/"

# List all files in the directory with full paths
ver_files <- list.files(ver_dir, full.names = TRUE)

# Function to process each file
grab_verified_data <- function(file_path) {
  # Extract the site name using the gsub method
  site_name <- gsub(paste0(ver_dir, "/"), "", file_path) %>% gsub("-.*", "", .)
  
  # Read the data and transform it
  single_site_param <- readRDS(file_path) %>%
    #exclude data that did not pass verification
    filter(!is.na(mean_verified))%>%
    select(DT_round, parameter, mean_verified) %>%
    mutate(site = site_name)
  
  return(single_site_param)
}

# Apply process_file function to each file and combine the results
sensor_data <- map_dfr(ver_files, grab_verified_data)%>%
  mutate(date = as.Date(DT_round, tz = "MST")) 

# wide_sensor_data <- sensor_data%>%
#   pivot_wider(names_from = parameter, values_from = mean_verified, id_cols = c(site, DT_round))%>%
#   #create date for match up
#         mutate(date = as.Date(DT_round, tz = "MST")) 



rm(ver_dir, ver_files, grab_verified_data, sensor_meta)
```




# Joining grab and sensor data

The current method looks at the verified data (erronous data removed already) and grabs the nearest four data points AFTER the grab sample was taken on the same day of sampling. If this data is not available (ie. Sonde was removed from field or was not recording ), data points preceeding the sampling are chosen.

If you wish to change the number of data points to grab, change the value of `data_points` in the chunk below. 

```{r}
####----Change n = 4 to the number of points you want to be selected----#####
data_points <- 4 

  merge_grab_sensor <- function( grab_dt, site_select){
    # convert to date to filter sensor data to only data on same date as sample
    grab_date = as.Date(grab_dt, tz = "MST")
    
    site_sonde_data <- sensor_data %>%
      filter(site == site_select & date == grab_date)
    
     if(nrow(site_sonde_data) == 0){
       return(NULL)
    }
      #get the available parameters for this site
      params <- unique(site_sonde_data$parameter)
      
      #fucntion to grab the nearest four datapoints to the grab sample (on the same day)
      grab_param_data <- function(sonde_data, parameter_arg){
        #filter to just the parameter of interest
        param_data <- sonde_data %>%
        filter(parameter == parameter_arg)%>%
          dplyr::arrange(DT_round)
     # grab data for each parameter after a sample was collected 
      after_data <- param_data %>%
        filter(DT_round >= grab_dt )%>%
        # grab the data 4 data points after a sample was collected
        dplyr::slice_head(n = data_points)
      
#grab data before grab sample time, this is to be used for when a sonde was removed, essentially reference data prior to the sonde being pulled rather than after
        before_data <- param_data %>%
          filter(DT_round < grab_dt )%>%
          # grab the data missing data points from after_data to make sure we have 4 data points (if there are already 4 then this will not grab any points before the sample collection)
####----Change 4 to the number of points you want to be selected----#####
          
          dplyr::slice_tail(n = (data_points- nrow(after_data)) )
        
        #bind data together
        select_param_data <- bind_rows(before_data, after_data)%>%
          mutate(site = site_select, grab_dt = grab_dt)
        return(select_param_data)
      }
      #grab sensor data for all available parameters
      select_sensor_data <- map(params, ~grab_param_data(sonde_data = site_sonde_data, 
                                                         parameter_arg = .x))%>%
        bind_rows()

#if no match up is found (should be caught earlier), return nothing
        if(nrow(select_sensor_data) == 0){
       return(NULL)
          
        }else{
    
    select_sensor_data%>%
      group_by(parameter)%>%
       # take the median of all sensor values
       summarise_if(is.numeric, median, na.rm = TRUE)%>%
         rename(median = mean_verified)%>%
         pivot_wider(names_from = parameter, values_from = median)%>%
       #add site and grab_dt back in
       mutate(site = site_select, grab_dt = grab_dt, 
              sonde_data_from = ifelse(grepl("virridy", site_select) == TRUE, "virridy", "csu or single sonde"))%>%
            #add all the sensor data used for future reference
            bind_cols(nest(select_sensor_data))
      }
    
  }

#sites with sensors at them   
sites <- c("JOEI", "CBRI", "CHD", 
           "PFAL", "SFM", "LBEA", "PENN", 
           "PBD", "lincoln", "timberline", "prospect","boxelder",  "archery", 
           "boxcreek", "springcreek")

#Only grab samples which are at sensor sites
lab_data_at_sensors <- most_recent_chem%>%
  #remove data before sensor data starts
  filter( site_code %in% sites, Year >= 2023)%>%
  #lower case to match sensor data
  mutate(site_code = tolower(site_code), 
         grab_dt = round_date(DT_mst, "15 minutes"))%>%
  #grab 
  select(site_code, grab_dt, Turbidity:SO4, Field_DO_mgL,Field_Cond_µS_cm,Field_Temp_C, flow_gauge_source, flow_gauge_id, hourly_q_cfs )%>%
  rename(lab_pH = pH, lab_turbidity = Turbidity)%>%
  #add in duplicate rows for when a site sampled has a virridy and csu sonde
  {bind_rows( {.}, {.} %>% filter(site_code %in% c("archery", "timberline", "prospect"))%>%
      mutate(site_code = paste0(site_code, " virridy"))
  )}




      
  correlated_df <- lab_data_at_sensors%>%
    #grab only dt and site from discrete samples for calculate avg fucn
    select(grab_dt, site_select = site_code)%>%
    # map over calc avg fucn
    pmap_dfr(., merge_grab_sensor)%>%
    #remove any rows that have no sensor data
    filter(rowSums(!is.na(select(., -grab_dt, -site, -sonde_data_from))) > 0)%>%
    # join back in lab data by site and grab dt
    left_join(lab_data_at_sensors, by = c("grab_dt", "site" = "site_code"))
  
  
  tidy_correlated_df <- correlated_df%>%
    #remove sites with no sensor data
     filter(map_lgl(data, ~!is.null(.)))%>%
    filter(!is.null(data))%>%
    select(site, grab_dt, data, everything())%>%
    arrange(site, grab_dt)
  
  

  rm(sites, correlated_df, merge_grab_sensor, lab_data_at_sensors)
```

# Save data to combined folder

```{r}

date <- format(Sys.Date(), "%Y%m%d")  
write_rds(tidy_correlated_df, paste0("data/combined/correlated_data_run_", date,".rds"))
  
```





